---
title: "Temporal distances"
author: "Makar Dorohuntsev"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,fig.width = 11)
```

### 1. Use function below to create data frame and fill it with calculated temporal distances

```{r}
#List of SNP names
load("~/Desktop/Data do projektu/Data/id by mafs.RData")
load("~/Desktop/Data do projektu/Data/cuted.RData")
#Take created list of frequencies from first step to get data and SNP names
 load("~/Desktop/Data do projektu/Data/filtred scores.RData")

Distance_count <- data.frame(filtred_id)
dist_2 <- 0
dist_5 <- 0
dist_9 <- 0
dist2_temp <- NULL
dist5_temp <- NULL
dist9_temp <- NULL
for (i in 1:102)
{
# Output is table with values from 0 to 1 where 0 is identical temporal frequencies and 1 is total domination of one. This table contains population 2,5 and 9 because only they have multiple temporal samples
  b <- filtred_id[i]
  temp <- freq_count[b]
  k = length(unlist(temp))
  if(k==22){j=0} else j=4
    # Pop 2
      j_2 <- unlist(temp,use.names = FALSE)[9] # read autumn allels
      w_2 <- unlist(temp,use.names = FALSE)[11] # read spring alleles
      if(k==33) # if there are 3 alleles, reduce them to 2 by adding first and third
      {j_2 <- unlist(temp,use.names = FALSE)[13] + unlist(temp,use.names = FALSE)[15] # sum of 1 and 3 allels
       w_2 <- unlist(temp,use.names = FALSE)[16] + unlist(temp,use.names = FALSE)[18]
        }
      ifelse((j_2+w_2)==0,0,mean_2 <- (j_2+w_2)/2)
      ifelse(j_2>w_2, p_2 <- j_2-w_2,p_2 <- w_2-j_2)
      ifelse((j_2+w_2)==0,0,dist_2 <- p_2/mean_2)
      dist2_temp <- c(dist2_temp,dist_2/2)
    # Pop 5
      j_5 <- unlist(temp,use.names = FALSE)[15] 
      w_5 <- unlist(temp,use.names = FALSE)[17]
      if(k==33)
      {j_2 <- unlist(temp,use.names = FALSE)[22] + unlist(temp,use.names = FALSE)[24]
        w_2 <- unlist(temp,use.names = FALSE)[25] + unlist(temp,use.names = FALSE)[27]
      }
      ifelse((j_5+w_5)==0,0,mean_5 <- (j_5+w_5)/2)
      ifelse(j_5>w_5, p_5 <- j_5-w_5,p_5 <- w_5-j_5)
      ifelse((j_5+w_5)==0,0,dist_5 <- p_5/mean_5)
      dist5_temp <- c(dist5_temp,dist_5/2)
      # Pop 9
      j_9 <- unlist(temp,use.names = FALSE)[19] 
      w_9 <- unlist(temp,use.names = FALSE)[21]
      if(k==33)
      {j_2 <- unlist(temp,use.names = FALSE)[28] + unlist(temp,use.names = FALSE)[30]
      w_2 <- unlist(temp,use.names = FALSE)[31] + unlist(temp,use.names = FALSE)[33]
      }
      ifelse((j_9+w_9)==0,0,mean_9 <- (j_9+w_9)/2)
      ifelse(j_9>w_9, p_9 <- j_9-w_9,p_9 <- w_9-j_9)
      ifelse((j_9+w_9)==0,0,dist_9 <- p_9/mean_9)
      dist9_temp <- c(dist9_temp,dist_9/2)
}

# Build data frame 
Distance_count["Pop 2"] <- dist2_temp
Distance_count["Pop 5"] <- dist5_temp
Distance_count["Pop 9"] <- dist9_temp

# Now we have data frame with calculated temporal distances of SNPs for every marker but rare markers with high distance mislead interpretation so they can be reduced by using MAF frequencies 
# optionaly options(digits = 12) to see more digits
Real_distance <- as.data.frame(Distance_count$filtred_id)
Real_distance$`Pop 2` <- Distance_count$`Pop 2`*cuted_test$Frequency
Real_distance$`Pop 5` <- Distance_count$`Pop 5`*cuted_test$Frequency
Real_distance$`Pop 9` <- Distance_count$`Pop 9`*cuted_test$Frequency

```

### 2. Creating plots from distance counts

Multiplot() function definition credits :

[http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page\_(ggplot2)/](http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/){.uri}

```{r include=FALSE}
  multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
    library(grid)
    
    # Make a list from the ... arguments and plotlist
    plots <- c(list(...), plotlist)
    
    numPlots = length(plots)
    
    # If layout is NULL, then use 'cols' to determine layout
    if (is.null(layout)) {
      # Make the panel
      # ncol: Number of columns of plots
      # nrow: Number of rows needed, calculated from # of cols
      layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                       ncol = cols, nrow = ceiling(numPlots/cols))
    }
    
    if (numPlots==1) {
      print(plots[[1]])
      
    } else {
      # Set up the page
      grid.newpage()
      pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
      
      # Make each plot, in the correct location
      for (i in 1:numPlots) {
        # Get the i,j matrix positions of the regions that contain this subplot
        matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
        
        print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                        layout.pos.col = matchidx$col))
      }
    }
  }
```

### Real distance values plot

This plot shows distribution of real distance values counted by taking relative distances and multiplying them by MAF frequencies. This allows to reduce low frequent and high different markers.

```{r}
library(ggplot2)
plot1 <- ggplot(Real_distance, aes(x=`Pop 2`)) + 
  geom_histogram(colour="black", fill="white",bins = 15)+
  labs(x="Temporal distance value",y="Number of markers")+
  ggtitle("Temporal distances for population 2")
plot2 <- ggplot(Real_distance, aes(x=`Pop 2`)) + 
  labs(x="Density of markers",y="Density value")+
  geom_density(alpha=.3, fill="#FF1626")+
  ggtitle("Density for population 2")

plot3 <- ggplot(Real_distance, aes(x=`Pop 5`)) + 
  geom_histogram(colour="black", fill="white",bins = 15)+
  labs(x="Temporal distance value",y="Number of markers")+
  ggtitle("Temporal distances for population 5")
plot4 <- ggplot(Real_distance, aes(x=`Pop 5`)) + 
  labs(x="Density of markers",y="Density value")+
  geom_density(alpha=.3, fill="#FF0666")+
  ggtitle("Density for population 5")

plot5 <- ggplot(Real_distance, aes(x=`Pop 9`)) + 
  geom_histogram(colour="black", fill="white",bins = 15)+
  labs(x="Temporal distance value",y="Number of markers")+
  ggtitle("Temporal distances for population 9")
plot6 <- ggplot(Real_distance, aes(x=`Pop 9`)) + 
  labs(x="Density of markers",y="Density value")+
  geom_density(alpha=.3, fill="#FF9")+
  ggtitle("Density for population 9")

multiplot(plot1,plot2,plot3,plot4,plot5,plot6,cols = 3)
```

### Percentage distances values

In this plot values represent range of differentiation in values from 0 to 1, where 0 identical temporal frequencies and 1 is presence of one and absence of other.

```{r echo=FALSE}
plot1 <- ggplot(Distance_count, aes(x=`Pop 2`)) + 
  geom_histogram(colour="black", fill="white",bins = 15)+
  labs(x="Percentage distance value",y="Number of markers")+
  ggtitle("Percentage distances for population 2")
plot2 <- ggplot(Distance_count, aes(x=`Pop 2`)) + 
  labs(x="Density of markers",y="Density value")+
  geom_density(alpha=.3, fill="#FF1626")+
  ggtitle("Density for population 2")

plot3 <- ggplot(Distance_count, aes(x=`Pop 5`)) + 
  geom_histogram(colour="black", fill="white",bins = 15)+
  labs(x="Percentage distance value",y="Number of markers")+
  ggtitle("Percentage distances for population 5")
plot4 <- ggplot(Distance_count, aes(x=`Pop 5`)) + 
  labs(x="Density of markers",y="Density value")+
  geom_density(alpha=.3, fill="#FF0666")+
  ggtitle("Density for population 5")

plot5 <- ggplot(Distance_count, aes(x=`Pop 9`)) + 
  geom_histogram(colour="black", fill="white",bins = 15)+
  labs(x="Percentage distance value",y="Number of markers")+
  ggtitle("Percentage distances for population 9")
plot6 <- ggplot(Distance_count
                , aes(x=`Pop 9`)) + 
  labs(x="Density of markers",y="Density value")+
  geom_density(alpha=.3, fill="#FF9")+
  ggtitle("Density for population 9")

multiplot(plot1,plot2,plot3,plot4,plot5,plot6,cols = 3)
```
